MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-23 paper.Rmd
Working on paper and refs.

Energy flows through society enable economic activity and facilitate human flourishing.
To understand economic growth and human well-being, 
the field of energy analysis evaluates energy flows
from the primary stage 
(coal, oil, natural gas) 
to the final stage
(energy purchased by consumers, such as refined petroleum and electricity)
and ultimately to the useful stage
(heat, mechanical drive, and light),
the last energy before energy services
(transport, thermal comfort, and illumination).
Societal exergy analysis (SEA), an extension of energy analysis,
quantifies energy flows as exergy, 
the physical work potential of energy.

@@

Historically, SEA has analyzed energy flows through specific countries
using spreadsheets, 
with data in varying and inconsistent formats 
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
The authors and others in the field
wanted to expand to cover all countries, but
the spreadsheet approach to data analysis would not scale.
We developed a series of `R` packages to enable 
analyses for all countries, as described in the following table.

@@

# Example usage

@@

>>>>

Energy flows through society enable economic activity and facilitate human flourishing.
To understand economic growth and human well-being, 
the field of energy analysis evaluates energy flows
from the primary stage 
(coal, oil, natural gas) 
to the final stage
(energy purchased by consumers, such as refined petroleum and electricity),
to the useful stage
(heat, mechanical drive, and light),
and sometimes to energy services
(transport, thermal comfort, and illumination).
Societal exergy analysis (SEA), an extension of energy analysis,
quantifies energy flows as exergy, 
the mechanical work potential of energy.

@@

Historically, SEA has analyzed energy flows through individual countries
using spreadsheets, 
with data in varying and inconsistent formats 
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa].
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis is not scalable.

To achieve scalable societal energy and exergy analysis, 
several sub-problems needed to be solved.
The first problem is the format for data storage. 
We use the Physical Supply-Use Table (PSUT) data structure,
a matrix approach to describing energy flows from 
extraction to
processing stages and, ultimately, 
to final demand [@Rocco:2016; @Guevara:2017; @Heun:2018].
A second problem is that `R` data fames.

It would be convenient to perform matrix mathematics in `R` data frames
as easily as scalar mathematics with the `dplyr` package from the `tidyverse`.
To do that, a way to embed matrices in a cell of a data frame is required.

Further, a means of manipulating row and column names

One problem is da

that different countries, 
and different years for a given country,
will have varying energy products and

To solve these problems and 
to meet the need for the SEA community,
we developed a suite of `R` packages.

# Design of `R` packages

to enable 
analyses for all countries, as described in the following table.

@@

# Example

@@

# Conclusion
